---
title: "Compute Percentage Changes of School Resources and Population"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compute Percentage Changes of School Resources and Population}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library(tibble)
library(dplyr)
library(tidyr)
library(readr)
library(kableExtra)

library(PrjCompPPTS)
```

In this file, we compute percentage changes in school resources and population. We interpolate when there is gaps in the data. 

For all variables, for each country/location, we compute: 

- percentage changes year by year
- percentage changes every 5, 10, 15, and 20 years. 


## Load and select

```{r}
# A. Load and select ----
ppts_easia_weuro_sel <- ppts_easia_weuro_world %>%
    select(country, area_in_country, year,
           contains("stats"))
```

## Transform data

```{r}
# B. Long to wide ----
ppts_easia_weuro_long <- ppts_easia_weuro_sel %>%
  pivot_longer(cols = starts_with('stats'),
               names_to = c('variable'),
               names_pattern = paste0("stats_(.*)"),
               values_to = "value")
str(ppts_easia_weuro_long)
# View(ppts_easia_weuro_long)


# C sort and group ----
ppts_easia_weuro_long <- ppts_easia_weuro_long %>%
    arrange(country, area_in_country, variable, year) %>%
    group_by(country, area_in_country, variable)
print(ppts_easia_weuro_long)
```

## Compute annual percentage changes

### Raw percentage changes

```{r}
# D. Annual percentage changes ----
# Compute these
# - annual: for all possible consecutive years
# - annual_interp1: annual based on consecutive year if possible, 
# when that is not possible, find closest years of available data, 
# and derive annual (considering compounding) growth rates

# D.1 annual
ppts_easia_weuro_long <- ppts_easia_weuro_long %>%
    mutate(pchg_yr1 = (value - lag(value)) / lag(value))
# View(ppts_easia_weuro_long)
```

### Interpolating 

```{r}
# D.2 annual_interp1
# D.2.1 compute linearly interpolated annual change
# create new sorting var, to take difference across years even if not
# consecutive
ppts_easia_weuro_interp1 <- ppts_easia_weuro_long %>%
    drop_na(value) %>%
# New sorting variable, +1 regardless of year gap
mutate(year_with_gap_ctr = row_number()) %>%
    arrange(country, area_in_country, variable, year_with_gap_ctr) %>%
    group_by(country, area_in_country, variable) %>%
# compute percentage change over span and years gap over span
mutate(pchg_span_interp1 = (value - lag(value)) / lag(value),
           span_yr = (year - lag(year))) %>%
# linear-interpolated annualized (with compounding) percentage change
mutate(pchg_yr1_interp1 = abs(1 + pchg_span_interp1) ^ (1 / span_yr) - 1) %>%
    select(country, area_in_country,
           year, variable, span_yr, pchg_yr1_interp1)

# D.2.2 expand to all interpolating years
# the pchg_yr1_interp1 variable is only shown at start year, 
# due to drop_na(value) earlier
# need to expand to all years
ppts_easia_weuro_interp1 <- ppts_easia_weuro_interp1 %>%
    ungroup() %>% mutate(span_yr_dup = span_yr) %>%
    drop_na(span_yr) %>%
    tidyr::uncount(span_yr) %>%
    group_by(country, area_in_country, variable, year) %>%
    mutate(year_adj = row_number() + year - span_yr_dup) %>%
    ungroup() %>%
    select(country, area_in_country,
           year_adj, variable, pchg_yr1_interp1) %>%
    rename(year = year_adj)
str(ppts_easia_weuro_interp1)
print(ppts_easia_weuro_interp1)
```

### Merge raw and interpolated results together

```{r}
# D.3 merge together
ppts_easia_weuro_long <- ppts_easia_weuro_long %>%
  left_join(ppts_easia_weuro_interp1,
            by = (c('country' = 'country',
                  'area_in_country' = 'area_in_country',
                  'year' = 'year',
                  'variable' = 'variable')))

str(ppts_easia_weuro_long)
print(ppts_easia_weuro_long)
```

## Generate percentage changes every 5, 10, 15, 20 years 

```{r}
# E. Generate cuts -----
ar_it_cut_types <- c(1, 2, 3, 4)
it_avg_type <- 1
it_cut_type <- 1
for (it_avg_type in c(1, 2)) {

  if (it_avg_type == 1) {
    svr_chg_var <- "pchg_yr1"
    # common var across all spans
    svr_chg_var_new <- "pchg_yrspan"
  } else if (it_avg_type == 2) {
    svr_chg_var <- "pchg_yr1_interp1"
    # common var across all spans
    svr_chg_var_new <- "pchg_yrspan_interp1"
  }

  # Add to stack the annual results
  ppts_easia_weuro_pchg <- ppts_easia_weuro_long %>%
        select(country, area_in_country,
                variable, year, one_of(svr_chg_var), value) %>%
        drop_na(one_of(svr_chg_var)) %>%
        rename(!!sym(svr_chg_var_new) := !!sym(svr_chg_var),
                year_bins = year) %>%
        mutate(year_bins = as.factor(year_bins)) %>%
        mutate(year_bins_type = 0) %>%
        ungroup()

  # Loop over cut types    
  for (it_cut_type in ar_it_cut_types) {

    # temp dataframe
    ppts_easia_weuro_long_cut <- ppts_easia_weuro_long %>%
            select(country, area_in_country,
                   variable, year,
                   one_of(svr_chg_var), value) %>%
            drop_na(one_of(svr_chg_var))

    # E.1 Cut types
    if (it_cut_type == 1) {
      ar_it_cuts <- seq(1940, 2020, length.out = 17)
    } else if (it_cut_type == 2) {
      ar_it_cuts <- seq(1940, 2020, length.out = 9)
    } else if (it_cut_type == 3) {
      ar_it_cuts <- seq(1945, 2020, length.out = 6)
    } else if (it_cut_type == 4) {
      # French no student data in 2019 
      ar_it_cuts <- seq(1940, 2020, length.out = 5)
    #   ar_it_cuts <- c(seq(1940, 2000, by = 20), 2019)
    }
    it_gap <- ar_it_cuts[2] - ar_it_cuts[1]

    ar_it_end_seg <- ar_it_cuts[2:length(ar_it_cuts)]
    ar_it_start_seg <- ar_it_end_seg - it_gap + 1
    ar_st_lab <- paste0(ar_it_start_seg, "-", ar_it_end_seg)
    # if (it_cut_type == 4) {
    #     ar_st_lab <- c(ar_st_lab[1:3], "2001-2019")
    # }

    # E.2 Generate new year groupings, consider only full-segments
    # consider only sub-segments with observations in all years
    ppts_easia_weuro_long_cut <- ppts_easia_weuro_long_cut %>%
            mutate(year_bins = cut(year,
                    breaks = ar_it_cuts,
                    labels = ar_st_lab,
                    right = TRUE)) %>%
            group_by(country, area_in_country,
                     variable, year_bins) %>%
            mutate(val_n_in_bin = n()) %>%
    # if gap is 5 years, all 5 years has annual chagne
    filter(val_n_in_bin == it_gap)

    # E.3 cumulative product
    ppts_easia_weuro_long_cut <- ppts_easia_weuro_long_cut %>%
            group_by(country, area_in_country,
                     variable, year_bins) %>%
            mutate(!!sym(svr_chg_var_new) :=
                   cumprod(1 + !!sym(svr_chg_var)) - 1)
    # View(ppts_easia_weuro_long_cut)

    # E.4, slices last row
    ppts_easia_weuro_long_cut <- ppts_easia_weuro_long_cut %>%
            slice(n()) %>%
            select(country, area_in_country,
                    variable, year_bins, one_of(svr_chg_var_new), value) %>%
            ungroup() %>%
            mutate(year_bins_type = it_cut_type)

    # E.5 Stack
    ppts_easia_weuro_pchg <- bind_rows(
            ppts_easia_weuro_pchg, ppts_easia_weuro_long_cut)

  }

  # export
  if (it_avg_type == 1) {
    ppts_easia_weuro_pchg_raw <- ppts_easia_weuro_pchg
  } else if (it_avg_type == 2) {
    ppts_easia_weuro_pchg_interp1 <- ppts_easia_weuro_pchg
  }
}
```
## Store to file 

We save both the raw file without interpolation and the file with gaps filled in by interpolation.

```{R}
# Write to CSV and write to rda
# write_csv(ppts_easia_weuro_pchg_raw, "data-raw/ppts_easia_weuro_pchg_raw.csv")
# usethis::use_data(ppts_easia_weuro_pchg_raw, overwrite = TRUE)
# write_csv(ppts_easia_weuro_pchg_interp1, "data-raw/ppts_easia_weuro_pchg_interp1.csv")
# usethis::use_data(ppts_easia_weuro_pchg_interp1, overwrite = TRUE)

ppts_easia_weuro_world_pchg_raw <- ppts_easia_weuro_pchg_raw
# write_csv(ppts_easia_weuro_world_pchg_raw, "data/ppts_easia_weuro_world_pchg_raw.csv")
usethis::use_data(ppts_easia_weuro_world_pchg_raw, overwrite = TRUE)
ppts_easia_weuro_world_pchg_interp1 <- ppts_easia_weuro_pchg_interp1
# write_csv(ppts_easia_weuro_world_pchg_interp1, "data/ppts_easia_weuro_world_pchg_interp1.csv")
usethis::use_data(ppts_easia_weuro_world_pchg_interp1, overwrite = TRUE)
```
